### 2020-08-11 abstract.tex
Introduction Stuff

ABSTRACT WILL BE REWRITTEN An application implemented using RxJS is unintuitive and time-consuming to debug. We will explore the practices software engineers apply during debugging and take a close look at the challenges they face in this workshop paper. On the way to propose solution concepts, we will validate previously identified struggles in an observational study and elaborate on previous research efforts in the field of reactive debugging.