### 2020-11-02 slides.tex
Switch to Overleaf

\documentclass[aspectratio=169]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tikz}

\usepackage{pgfpages}
\setbeameroption{show notes on second screen}
% \pgfpagesuselayout{2 on 1}[a4paper,border shrink=10mm]

\usetheme[darkmode, nofooterlogo]{pureminimalistic}
% \renewcommand{\pageword}{}
% \renewcommand{\logoheader}{\vspace{1.5em}}
\usepackage[backend=biber, doi=false, maxbibnames=2, maxcitenames=2, style=numeric, sorting=none, url=false, eprint=false]{biblatex}
\addbibresource{bibliography.bib}
% this makes it possible to add backup slides, without counting them
\usepackage{appendixnumberbeamer}
\renewcommand{\appendixname}{\texorpdfstring{\translate{appendix}}{appendix}}

\title[Debugging RxJS-based Applications]{Getting rid of the \texttt{console.log}:\\Debugging RxJS-based Applications}
\author{Manuel Alabor}
\institute{Eastern Switzerland University of Applied Sciences}
\date{\today}

\begin{document}

\maketitle

\note[itemize]{
	\item Welcome
	\item Debugging of RxJS-based Applications
}

\section*{Introduction}

\begin{frame}{Who Am I?}
	\begin{vfilleditems}
		\item Foobar\bigskip
		\item Foobar\bigskip
		\item Foobar\bigskip
		\item Foobar\bigskip
	\end{vfilleditems}
\end{frame}

% \begin{frame}[focus]
% 	\includegraphics[width=\textwidth/3]{figures/rxjs-logo.png}
% 	\\
% 	\vspace{0.5cm}
% 	\Huge{RxJS}
% \end{frame}

\note[itemize] {
	\item
}

\section*{Introduction}

\begin{frame}{}

\end{frame}

\note[itemize] {
	\item
}

\begin{frame}{}

\end{frame}

\note[itemize] {
	\item
}

\section{Interviews and War Stories}

\begin{frame}{Interviews}

\end{frame}

\note[itemize] {
	\item
}

\begin{frame}{War Stories}

\end{frame}

\note[itemize] {
	\item
}

\begin{frame}{Results}

\end{frame}

\note[itemize] {
	\item
}

\section{Observational Study}
\begin{frame}{Hypothesis}

\end{frame}

\note[itemize] {
	\item
}

\begin{frame}{Design}

\end{frame}

\note[itemize] {
	\item
}

\begin{frame}{Problem Example}

\end{frame}

\note[itemize] {
	\item
}

\begin{frame}{Results}

\end{frame}

\note[itemize] {
	\item
}

\section{Conclusion}
\begin{frame}{}

\end{frame}

\note[itemize] {
	\item
}

\begin{frame}{Future Work}

\end{frame}

\note[itemize] {
	\item
}

\begin{frame}
	Q \& A
\end{frame}

\appendix
\begin{frame}[allowframebreaks]{References}
	\bibliographystyle{plain}
	\bibliography{bibliography}
\end{frame}

\end{document}